---
title: "Marketing Performance & Growth Efficiency at trivago (2023–2025)"
author: "Aleksei Prishchepo"
date: "2026-01-11"
fig-format: svg
bibliography: references.bib
csl: harvard-cite-them-right.csl
css: styles.css
format:
  pdf:
    papersize: a4
    toc: true
    toc-depth: 2
    fig-pos: 'H'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(ggplot2)
library(dplyr)
library(tidyr)
library(kableExtra)
```

## Executive Summary

**trivago** is executing a multi-year strategic turnaround aimed to
ignite its brand by shifting from short-term profit maximization toward
long-term brand reinvestment. Since late 2024 company returned to
year-over-year revenue growth.

### Key insights

-   Advertising investment accelerated materially in 2025 (+24% YoY in
    Q1), while Global ROAS decreased from \~148% in 2023 to \~118% in Q2
    2025, largely due to increased branded spend.

-   Developed Europe continues to be the most efficient region, while
    Americas rebounded strongly in Q2 2025.

-   Growth quality has improved through higher branded traffic,
    increased user loyalty (logged-in users now contribute 20% of
    referral revenue in comparison to 10% in 2023), and meaningful
    conversion gains driven by product and AI initiatives.

```{r tbl-key-metrics-2024-2025}
#| tbl-cap: Key Metrics Overview (2024-2025)

dataset <- read.csv("data/key-metrics-2024-2025.csv")

dataset <- janitor::clean_names(dataset)
dataset <- rename(
  dataset,
  c(
    "Metric" = "metric",
    "Latest Data (2024-2025)" = "latest_data_2024_2025"
  )
)
dataset |>
  kable(format = "latex", longtable = FALSE, booktabs = TRUE) |>
  kable_styling(bootstrap_options = "striped", full_width = TRUE)
```

## Business and Data Context

trivago operates a performance-driven metasearch model where marketing
efficiency is a primary strategic lever. Referral revenue is generated
predominantly through CPC-based traffic acquisition, making Return on
Advertising Spend (ROAS) a central decision metric rather than a purely
operational KPI.

The analysis is based on aggregated financial, marketing, and
operational disclosures covering 2023–2025. Data limitations include:

-   Geographic segmentation rather than country-level or channel-level
    granularity

-   A mix of quantitative metrics and qualitative management commentary

-   Absence of user-level or campaign-level performance data

Given these constraints, the report focuses on directional insights,
efficiency trends, and strategic implications, rather than tactical
optimization.

## Marketing Investment and Efficiency Outcomes

### Advertising Strategy

In 2025, trivago materially increased advertising investment to rebuild
a “branded visitor baseline”, which management views as essential for
sustainable growth.

::: {#tbl-advertising-spend-by-segment tbl-pos="h" layout-ncol="1" tbl-cap="Advertising spend by segment (in million EUR)"}
```{r advertising-spend-by-segment}
dataset <- read.csv("data/advertising-spend-by-segment-m-eur.csv")
dataset <- janitor::clean_names(dataset)
dataset <- dataset |> filter(grepl("2023|2024|2025", period))

dataset <- rename(
  dataset,
  c(
    "Period" = "period",
    "Americas" = "americas",
    "Developed Europe" = "developed_europe",
    "Rest of World" = "rest_of_world",
    "Total Advertising Spend" = "total_global"
  )
)

dataset |>
  kable(format = "latex", longtable = FALSE, booktabs = TRUE) |>
  kable_styling()
```
:::

In Q1 2025, total Advertising Spend rose by 24% YoY, reaching €104.5
million (see @tbl-advertising-spend-by-segment).

Incremental investment has been disproportionately allocated toward
branded channels (direct visits and brand-related search), rather than
generic performance marketing auctions.

Generative AI has been used to scale marketing execution, including
AI-localized TV advertising across more than ten languages using
creative adaptation.

```{r fig-advertising-spend-vs-referral-revenue, fig.width=6, fig.height=2.5}
#| fig-cap: Advertising Spend vs. Referral Revenue (2022-2024)

revenue <- read.csv("data/referral-revenue-by-segment-m-eur.csv")
revenue <- janitor::clean_names(revenue)

ad_spend <- read.csv("data/advertising-spend-by-segment-m-eur.csv")
ad_spend <- janitor::clean_names(ad_spend)

dataset <- revenue |> inner_join(ad_spend, by = "period", suffix = c(".rev", ".ad"))
dataset <- rename(
  dataset,
  c(
    "Period" = "period",
    "Total Referral Revenue" = "total_referral_revenue",
    "Total Advertising Spend" = "total_global"
  )
)
dataset <- dataset |> filter(grepl("2022|2023|2024", Period))

dataset |> ggplot() +
  geom_line(aes(x = Period, y = `Total Referral Revenue`, group = 1, color = "Total Referral Revenue")) +
  geom_line(aes(x = Period, y = `Total Advertising Spend`, group = 1, color = "Total Advertising Spend")) +
  scale_y_continuous(limits = c(0.0, NA)) +
  labs(
    x = NULL,
    y = "Amount (Million EUR)",
    color = "Metric"
  ) +
  theme_minimal()

```

### Marketing Efficiency and ROAS

Despite accelerated spend, marketing efficiency has remained stable at
the global level.

Global ROAS reached 119.0% in Q2 2025 vs. 132.10% in the prior year.

```{r fig-roas-trend-by-segment, fig.width=6, fig.height=2.5}
#| fig-cap: ROAS trend by geographic segment

dataset <- read.csv("data/roas-by-segment-percent.csv")
dataset <- janitor::clean_names(dataset)

dataset <- dataset |> filter(grepl("2023|2024|2025", period))

dataset |> ggplot() +
  geom_line(aes(x = period, y = developed_europe, group = 1, color = "Developed Europe")) +
  geom_line(aes(x = period, y = americas, group = 1, color = "Americas")) +
  geom_line(aes(x = period, y = rest_of_world, group = 1, color = "Rest of World")) +
  geom_line(aes(x = period, y = global_total, group = 1, color = "Global ROAS"), size = 1.2) +
  # start y from 0
  scale_y_continuous(limits = c(0.0, NA)) +
  labs(
    # title = "ROAS Trend by Geographic Segment",
    x = NULL,
    y = "ROAS (%)",
    color = "Segment"
  ) +
  theme_minimal()
```

Developed Europe remains the most efficient region, with ROAS of 134%,
benefiting from strong returns on earlier brand investments (see
@tbl-roas-contribution-by-segment).

::: {#tbl-roas-contribution-by-segment tbl-pos="h" layout-ncol="1" tbl-cap="ROAS contribution by segment (in million EUR)"}
```{r roas-contribution-by-segment}
dataset <- read.csv("data/roas-contibution-by-segment-m-eur.csv")
dataset <- janitor::clean_names(dataset)
dataset <- dataset |> filter(grepl("2023|2024|2025", period))
dataset <- rename(
  dataset,
  c(
    "Period" = "period",
    "Americas" = "americas", "Developed Europe" = "developed_europe",
    "Rest of World" = "rest_of_world",
    "Total ROAS Contribution" = "total_global"
  )
)
dataset |>
  kable(format = "latex", longtable = FALSE, booktabs = TRUE) |>
  kable_styling()
```
:::

In the Rest of World, referral revenue increased 44% YoY in Q1 2025,
reflecting high growth elasticity during post-COVID recovery.

In the first half of 2025, trivago generated €41.0 million in ROAS
contribution, representing referral revenue remaining after advertising
spend.

Source: @trivago_annual_website, @trivago_q1_2025_review,
@refinitiv2025, @globenewswire2025_q1, @globenewswire2025_q2

## Growth Drivers

### Traffic Composition and Demand Quality

A core driver of improved growth efficiency is the shift in traffic
composition toward higher-quality demand sources.

Branded channel traffic achieved double-digit growth across all segments
in 2025.

This shift reduces exposure to increasingly competitive and volatile
performance marketing auctions and supports more stable long-term ROAS.

+---------------+------------------------+--------------------------+
| Period        | Branded Channel        | Performance Marketing    |
|               | Traffic & Strategy     | Channel & Strategy       |
+===============+========================+==========================+
| 2019          | Characterized by high  | Significant investment   |
|               | brand awareness and    | in high-volume keyword   |
|               | **extensive TV         | auctions.                |
|               | marketing**.           |                          |
+---------------+------------------------+--------------------------+
| 2020 - 2022   | **Brand spend ceased   | Primary source of        |
|               | almost completely** in | traffic, but managed     |
|               | 2020 to preserve       | with **high ROAS         |
|               | liquidity. Resumed at  | targets** to maintain    |
|               | highly reduced levels  | profitability during low |
|               | in late 2021.          | demand.                  |
+---------------+------------------------+--------------------------+
| 2023          | Multi-year strategy to | Experienced **increased  |
|               | "reignite" the brand   | volatility** and         |
|               | announced. New global  | headwinds due to Google  |
|               | creative campaign      | ad format changes.       |
|               | launched in December.  |                          |
+---------------+------------------------+--------------------------+
| 2024          | Branded traffic growth | Faced higher competition |
|               | reached double-digit   | and rising costs in      |
|               | levels in Developed    | keyword auctions.        |
|               | Europe and Americas.   |                          |
|               | Brand spend increased, |                          |
|               | driving a 7% rise in   |                          |
|               | total ad costs.        |                          |
+---------------+------------------------+--------------------------+
| 2025 H1       | **20% of total         | Continued disciplined    |
|               | Referral Revenue**     | approach; performance    |
|               | generated by logged-in | marketing spend reduced  |
|               | members, double the    | in some segments (e.g.,  |
|               | share of 2023. Higher  | Europe) to prioritize    |
|               | proportional           | brand elasticity.        |
|               | investment in brand    |                          |
|               | vs. performance.       |                          |
+---------------+------------------------+--------------------------+

: Traffic mix evolution: Branded vs. Performance

### User Loyalty and Retention

User loyalty has emerged as an increasingly important contributor to
revenue quality.

In Q2 2025, logged-in users generated 20% of total referral revenue,
doubling their contribution since 2023.

This trend reflects improved retention and repeat usage, increasing the
lifetime value of acquired users and lowering effective acquisition
costs.

### Conversion Efficiency Improvements

Conversion gains have amplified the returns from marketing investment.

trivago’s internal booking conversion index improved materially in 2025,
with efficiency gains of approximately 35%.

“Project Trinity” simplified the price comparison experience by
restructuring how rates are grouped and displayed, leading to higher
user confidence and conversion.

AI-generated hotel highlights have expanded to 350,000+ hotels, reducing
research friction for users.

### Marketplace and Product Enablers

In parallel with traffic and conversion improvements, trivago has
implemented structural changes to stabilize and scale its marketplace.

More than 100 partners have transitioned to a CPA-based bidding model,
reducing auction volatility.

Transaction-based pricing has doubled its share of marketplace revenue
since 2023.

These changes improve predictability of returns for both advertisers and
trivago, supporting sustainable marketing scale.

### Acquisition of Holisto Limited

trivago's acquisition of Holisto Limited, an AI-driven travel technology
platform and hotel rate aggregator, was executed to enhance its
technical capabilities and revenue streams.

The primary implication of this acquisition is the accelerated rollout
of "trivago Book & Go", a white-label booking engine that allows
travelers to complete reservations through a seamless, trivago-branded
funnel.

Financially, management expects Holisto to contribute a low double-digit
million euro increase to consolidated total revenue in 2025 while
operating near a breakeven level. Strategically, the integration enables
trivago to shift more of its marketplace toward a transaction-based
(CPA) model.

Source: @trivago_annual_website, @globenewswire2025_q1,
@globenewswire2025_q2, @refinitiv2025, @mirai2024, @dcfmodeling2025,
@ibisworld2025, @stockanalysis2025

## Risks, Constraints and External Factors

Despite renewed growth momentum, trivago continues to face several
structural challenges:

-   Platform dependency: Changes to search engine advertising formats
    negatively impacted traffic volumes throughout 2023 and early 2024.

-   Advertiser concentration: Expedia Group and Booking Holdings account
    for approximately 75% of referral revenue, creating dependency risk
    (see @fig-advertiser-concentration).

```{r fig-advertiser-concentration, fig.width=6, fig.height=2.5}
#| fig-cap: Advertiser Concentration

dataset <- read.csv("data/advertiser-concentration-percent.csv")
dataset <- janitor::clean_names(dataset)
dataset <- dataset |> filter(grepl("2023|2024|2025", period))
dataset <- rename(
  dataset,
  c(
    "Period" = "period",
    "Expedia Group" = "expedia_group_share",
    "Booking Holdings" = "booking_holdings_share"
  )
)

dataset <- dataset |> pivot_longer(
  cols = c("Expedia Group", "Booking Holdings"),
  names_to = "Advertiser",
  values_to = "Share"
)

dataset |>
  ggplot(aes(
    x = Period, y = Share, fill = Advertiser,
    group = Advertiser, colour = Advertiser
  )) +
  geom_area(position = "stack", alpha = 0.5) +
  geom_line(position = "stack", size = 0.25) +
  scale_y_continuous(limits = c(0, 100), expand = c(0, 0)) +
  labs(
    x = NULL,
    y = "Share of Referral Revenue (%)",
    fill = "Advertiser"
  ) +
  theme_minimal()
```

-   Regulatory environment: The EU Digital Markets Act (DMA) has
    weakened Google’s hotel products, creating short-term traffic
    tailwinds in Europe, though the durability of these effects remains
    uncertain.

Refer to @tbl-swot-matrix for a comprehensive SWOT analysis.

::: {#tbl-swot-matrix tbl-pos="h" layout-ncol="1" tbl-cap="trivago SWOT Matrix (2025)"}
+---------------------------------+----------------------------------+
| **STRENGTHS**                   | **WEAKNESSES**                   |
+=================================+==================================+
| -   **Global Scale:** Access to | -   **Advertiser                 |
|     **5.0 million+              |     Concentration:** Heavy       |
|     accommodations** in over    |     reliance on **Expedia        |
|     190 countries.              |     Group** (35-37%) and         |
|                                 |     **Booking Holdings**         |
| -   **Brand Equity:** High      |     (37-40%) for the majority of |
|     global recognition with the |     revenue.                     |
|     iconic **"Hotel? trivago"** |                                  |
|     tagline.                    | -   **High Acquisition Costs:**  |
|                                 |     Advertising spend typically  |
| -   **AI Innovation:** Advanced |     consumes **over 80% of total |
|     proprietary matching        |     revenue**.                   |
|     algorithms and **AI Smart   |                                  |
|     Search** capabilities.      | -   **Financial Volatility:**    |
|                                 |     History of **net losses**    |
| -   **Direct Traffic Growth:**  |     and susceptibility to rapid  |
|     Strong branded channel      |     changes in advertiser        |
|     traffic; **20% of referral  |     bidding strategies.          |
|     revenue** now comes from    |                                  |
|     logged-in members.          | -   **Mobile Conversion Gap:**   |
|                                 |     Visitors via mobile devices  |
| -   **Operational Agility:**    |     traditionally **convert at   |
|     Ability to **rapidly adjust |     lower rates** than desktop   |
|     marketing spend** in        |     users.                       |
|     response to market          |                                  |
|     volatility.                 | -   **Niche Focus:** Historical  |
|                                 |     focus almost exclusively on  |
| -   **Strategic M&A:**          |     accommodations compared to   |
|     Acquisition of **Holisto    |     "all-in-one" travel OTAs.    |
|     Ltd** to provide            |                                  |
|     white-label booking and AI  |                                  |
|     rate aggregation.           |                                  |
+---------------------------------+----------------------------------+

+---------------------------------+----------------------------------+
| **OPPORTUNITIES**               | **THREADS**                      |
+=================================+==================================+
| -   **Branded Baseline          | -   **Search Engine Dominance:** |
|     Rebuilding:** Multi-year    |     **Google Hotel Ads**         |
|     strategy to increase        |     competing directly by        |
|     **direct traffic** and      |     promoting its own products   |
|     reduce reliance on paid     |     at the expense of organic    |
|     search.                     |     search.                      |
|                                 |                                  |
| -   **Vertical                  | -   **Macroeconomic Pressures:** |
|     Diversification:**          |     **Inflation and rising       |
|     Potential to expand into    |     costs** potentially          |
|     **flights, car rentals, and |     dampening consumer           |
|     activities** through        |     discretionary spending on    |
|     partnerships.               |     travel.                      |
|                                 |                                  |
| -   **CPA Model Transition:**   | -   **AI Disruption:** Rise of   |
|     Moving partners to a        |     **AI hyperscalers**          |
|     transaction-based model to  |     (ChatGPT, Gemini)            |
|     reduce auction volatility.  |     potentially changing how     |
|                                 |     travelers research and book  |
| -   **EU Regulatory Shifts:**   |     travel.                      |
|     Gaining market share as the |                                  |
|     **Digital Markets Act       | -   **Regulatory Scrutiny:**     |
|     (DMA)** restricts Google's  |     Persistent investigations    |
|     ability to self-preference  |     into **pricing transparency  |
|     its own hotel products in   |     and ranking algorithms** by  |
|     the EU.                     |     global competition.          |
+---------------------------------+----------------------------------+
:::

## Strategic Implications

The strategic implications center on a multi-year turnaround effort that
prioritizes long-term brand health and technological defensive measures
over short-term profit maximization. The primary strategic implications
include:

-   **Reduced Dependency on Paid Search:** A core weakness is trivago's
    reliance on performance marketing channels (generic search engine
    keywords), where high competition drives up acquisition costs.

-   **Prioritizing Brand Investment:** Management is shifting strategy
    to "reignite" the brand through massive reinvestment ad campaigns
    (e.g., the Jürgen Klopp campaign). The goal is to build a baseline
    of **direct, branded traffic** which is more loyal and
    higher-converting than paid search leads.

-   **Acceptance of Short-term Losses:** This strategy implies a
    deliberate decision to prioritize long-term direct user loyalty,
    even if it negatively impacts near-term profitability or leads to
    operating losses.

-   **Hedge against Hyperscalers:** trivago views the integration of AI
    (such as **AI Smart Search** and personalized ranking) as a
    necessary defense against AI hyperscalers like ChatGPT and Gemini,
    which threaten to disrupt how travelers research and book.

-   **Shift to Transaction-Based Revenue:** To mitigate the volatility
    caused by a few large advertisers changing their cost-per-click
    (CPC) bidding strategies, trivago is aggressively transitioning
    partners to a **Cost-Per-Acquisition (CPA)** model.

-   **Empowering Smaller Partners:** The acquisition of **Holisto Ltd**
    enables the rollout of "**trivago Book & Go**," a white-label
    booking engine. This allows smaller partners without strong brands
    to benefit from a trusted, trivago-branded checkout, thereby
    reducing reliance on the "Big Two" (Expedia and Booking Holdings).

-   **EU Market Share Gains:** The implementation of the **Digital
    Markets Act (DMA)** in the EU restricts Google's ability to
    self-preference its own hotel products. This presents a strategic
    opportunity for trivago to regain metasearch market share as Google
    Hotels loses visibility in the region.

-   **Fostering Membership and Loyalty**: trivago is increasingly
    focusing on user retention by offering exclusive deals and price
    alerts for members. The fact that **20% of Referral Revenue** now
    stems from logged-in users indicates a successful shift toward a
    more predictable and engaged user base.

{{< pagebreak >}}

## References

::: {#refs}
:::
